---
title: "Additional Networks"
author: "Steele H. Valenzuela & Jonathan Chiang"
date: "March 24, 2015"
output: html_document
---

#### Introduction

Echoing the introduction of the *README.md*, this project is the collection of all FBS college football games in the 2012 season. I am interested in breaking from the norm of the traditional boxscore and displaying schedules in a manner of simple and complex networks.

Although the networks are more on the complex and cluttered side of the aisle, what we have learned from this project is data scraping, cleaning, and manipulating into a 'workable' format in order to display data in a network. 

#### Libraries and Data

For a quick run through of this project, just follow the directions. The libraries we used are the `dplyr` and `networkD3` packages. If you have never loaded them in your library, please install them first.
```{r, message = FALSE}
library(dplyr)
library(networkD3)
```

Next, download these files from the CSVs folder (in my repo) and load them into your `R` working directory with the following commands. 
```{r}
cleansched7 <- tbl_df(read.csv("cleansched7.csv", header = T)[-1])
teamchr <- tbl_df(read.csv("teamchr.csv", header = TRUE)[-1])
```

In order to familiarize yourself with the tables/data frames, simply use the `slice` function from the `dplyr` package or simply type the name of the data frame into the command line and press enter. The first 10 rows of each table/data frame will display various information. The data frame **teamchr** will display the team name, abbreviation, and conference.

The first 25 rows of the data frame **cleansched7** display the pre-season top 25 rankings while the last 25 rows display the post-season top 25 rankings. Each row represents a team, their information such as conference and abbreviation, followed by the outcome of their game during that week. The next line should follow with their opponent and similar information.

#### Network of Entire Season

First, follow these commands to prepare the data for the networks. You may run this entire chunk and *skip* to the next chunk in order to view the said network as this piece is a bit *technical*
```{r}
odd_index <- seq(27, 1705, 2) # winner/source
even_index <- seq(26, 1704, 2) # loser/target

Source <- cleansched7$id[odd_index] - 1 # Source and Target must start at zero. A bit confusing.
Target <- cleansched7$id[even_index] - 1 # Source and Target must start at zero. A bit confusing.
value <- cleansched7$ptdiff[odd_index] # winner's pt differential for how wide the links are

name <- as.factor(teamchr$name) # names need to be factored
group <- as.integer(as.factor(teamchr$conference)) # groups must be integers

schedLinks <- tbl_df(data.frame(Source, Target, value))
schedNodes <- tbl_df(data.frame(name, group))
```

The following command will display a simple network, as the function suggests. Please note, it's quite messy. 
```{r Entire Simple Network}
simpleNetwork(Data = schedLinks, Source = "Source", Target = "Target")
```

The following command contains a few more options to spice up the network. Once again, it's a bit messy and may slow your computer down. 
```{r Entire Force Network}
forceNetwork(Links = schedLinks, Nodes = schedNodes, 
             Source = "Source", Target = "Target", 
             Value = "value", NodeID = "name", Group = "group", 
             fontsize = 15, opacity = .8)
```


#### Non-Conference Network

A majority of the non-conference games are played within the first four weeks of the season. I have subsetted the data to follow suit. You will see floating nodes that are not connected. That simply means that they are scheduled for later in the season. 
```{r}
ncf <- cleansched7 %>% filter(week <= 4 & week != 0) %>% select(id, ptdiff)

ncf_even_index <- seq(2, 552, 2) # winner/Source
ncf_odd_index <- seq(1, 551, 2) # loser/Target

ncfSource <- ncf$id[ncf_even_index] - 1; ncfTarget <- ncf$id[ncf_odd_index] - 1 
ncfValue <- ncf$ptdiff[ncf_even_index]
ncfLinks <- tbl_df(data.frame(ncfSource, ncfTarget, ncfValue))

forceNetwork(Links = ncfLinks, Nodes = schedNodes, 
             Source = "ncfSource", Target = "ncfTarget", Value = "ncfValue", 
             NodeID = "name", Group = "group", 
             fontsize = 15, opacity = 1)
```

### Conference Network

Conference games typically occur after week 4 until week 15, where the final regular season game is played between Army and Navy. That is displayed below. You'll notice the clusters of in-conference play but you'll also notice the few games that are not played within conference.
```{r Conference Network}
# last game before bowl games is Army vs Navy, which is lone game in Week 15
cf <- cleansched7 %>% filter(week > 4 & week <= 15) %>% select(id, ptdiff)

cf_even_index <- seq(2, 1058, 2) # winner/Source
cf_odd_index <- seq(1, 1057, 2) # loser/Target

cfSource <- cf$id[cf_even_index] - 1; cfTarget <- cf$id[cf_odd_index] - 1
cfValue <- cf$ptdiff[cf_even_index]
cfLinks <- tbl_df(data.frame(cfSource, cfTarget, cfValue))

forceNetwork(Links = cfLinks, Nodes = schedNodes, 
             Source = "cfSource", Target = "cfTarget", Value = "cfValue", 
             NodeID = "name", Group = "group", 
             fontsize = 15, opacity = 1, charge = -100)
```

If you don't know what you're looking for, it's a bit tricky. The salmon/pink colored nodes of **Army**, **Navy**, **Brigham Young**, and **Notre Dame** are quite interesting as they are independent of any conference and schedule whoever they want. Try to find **Notre Dame** and their list of opponents to see how many networks they are connected to.

### Display of conferences and their respective teams

Here we have transformed the data frame **teamchr** into its own network. 
```{r Within Conference Network}
# re-factored data frame to display names
confNodes <- tbl_df(data.frame(name = as.factor(teamchr$name), 
                               group = as.factor(teamchr$conference)))

# charge set to -100 to instead of default -200. Easier to fit all teams in one frame
simpleNetwork(Data = confNodes, Source = "name", Target = "group", 
              fontSize = 15, opacity = 1, charge = -100)
```
You will notice the many teams that make up the **FCS** conference as well as a few of the smaller ones, such as the **Independents** and **Division II** opponents. 